---
title: "p8105_hw3_rw2708"
author: "Ran Wang"
date: "10/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

#Problem 1
```{r instacart}
data("instacart") 
instacart <- instacart %>% 
 select(user_id, order_id, order_number, product_id, product_name, order_dow, everything())

```
Description: the dataset contains the online grocery purchase information for instacart. The dataset has `r nrow(instacart)` of observations and `r ncol(instacart)` of variables. The key variables in this dataset are `r colnames(instacart)[1:2]` that identify each customer and their order id, `r colnames(instacart)[3:5]` indicate the quantity and name of products that each customer purchased. For example,


```{r instacart}
 count_aisle <- instacart %>%
  arrange(aisle_id,aisle) %>% 
  group_by(aisle_id,aisle) %>% 
  count(aisle, name = "count_aisle",sort = TRUE)
```
There are `r nrow(count_aisle)` aisles and the `r count_aisle[1,2]` are the are the most items ordered from. 


```{r plot_aisle}
plot_aisle <- count_aisle %>% 
  filter(count_aisle >= 10000) %>% 
   ggplot(aes(x = aisle, y = count_aisle)) + 
  geom_point(aes(color = count_aisle)) +
   labs(title = "aisle plot",x = "aisle id", y = "counts of aisle")

plot_aisle
```

```{r question3}
table_1 =  
  instacart %>% 
  arrange(aisle_id, aisle) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle,product_name) %>% 
  summarize(count = n()) %>% 
  mutate(pop_rank=dense_rank(desc(count))) %>% 
           filter(pop_rank < 4) %>% 
  arrange(pop_rank) %>% 
  pivot_wider(
  names_from = "pop_rank", 
  values_from = c("product_name", "count")) %>% 
  rename("Rank_1"=product_name_1,"Rank_2"=product_name_2,"Rank_3"=product_name_3,"times_rank1"=count_1,"times_rank2"=count_2,"times_rank3"=count_3) %>% 
  knitr::kable(digits = 1)
  
```

```{r question4}
table_2 =  
  instacart %>% 
  arrange(product_name,order_hour_of_day,order_dow) %>% 
  filter(product_name=="Pink Lady Apples"| product_name =="Coffee Ice Cream") %>% 
  group_by(product_name,order_dow) %>% 
  summarize(mean_day=mean(order_hour_of_day)) %>% 
  pivot_wider(
    id_cols = "order_dow",
  names_from = "product_name", 
  values_from = "mean_day")%>% 
  rename("Day of Week"=order_dow) %>% 
  knitr::kable(digits = 1)
  
```

#Problem 2
```{r BRFSS}
library(p8105.datasets)
 



```



  


